Diffusive Homeostasis in a Self-Organizing Recurrent Neural Network \\ \begin{large} Spatially dependent Interaction as a Determinant of Neural Activity and Plasticity \end{large}

### 2016-12-14 draft.tex
.

Obviously, the appropriate choice of $NO_0$ is crucial for the goal of achieving and maintaining a certain level of activity. However, one cannot directly set a parameter of the model to the desired population activity, as it was the case for canonical intrinsic homeostasis. Rather, one needs to determine the average concentration \textit{associated} with the desired activity and set it as a target concentration. Though it is possible to derive this relation in an analytic fashion, for practical purposes of the simulation, we let the system run with the previous homeostatic mechanism, still solving equation \eqref{Ca_dyn}-\eqref{NO_dyn} until a steady mean over the concentrations at the neurons' positions was reached. This mean was then set to be the target concentration and we switched to diffusive homeostasis. Table \ref{Params_IP} summarizes the choice of parameters that where introduced in this section.
\begin{table}
\begin{tabular}{|l|l|}
\hline
\textbf{parameter} & \textbf{value} \\
\hline
$\mathrm{r_{IP}}$ & $\mathrm{3\;Hz}$ \\
\hline
$\mathrm{\eta_{IP}}$ & $\mathrm{0.1\;mV}$ \\
\hline
$\mathrm{Ca^{2+}_{spike}}$ & 1 \\ \hline
$\mathrm{\tau_{Ca^{2+}}}$ &  $\mathrm{10\;ms}$ \\
\hline
$\mathrm{\tau_{nNOS}}$ & $\mathrm{100\;ms}$ \\
\hline
$\mathrm{D}$ & default: 10000 $\mathrm{\mu m^2 s^{-1}}$ \\
\hline 
$\mathrm{\lambda}$ & $\mathrm{0.1\;s^{-1}}$ \\
\hline
$\mathrm{\tau_{V_t}}$ & see section \ref{activ_analys} \\
\hline
\end{tabular}
\caption{Parameters of homeostatic intrinsic plasticity.}
\label{Params_IP}
\end{table}

@@

\subsection{Activity Analysis} \label{activ_analys}
As a first attempt, we set the time constant of threshold adaptation in the diffusive homeostasis to $\mathrm{2500\, ms}$, as given by \cite{Sweeney_Paper}. Switching between non-diffusive/diffusive homeostasis after 750 seconds, Figure \ref{full_sim_osci} shows the resulting dynamics of the population activity of the excitatory and inhibitory group, the average $NO$ concentration at the excitatory neurons' positions and the average firing threshold within the exc. group (inh. threshold was fixed, see table \ref{LIF_neuron_params}). Both homeostatic mechanisms managed to keep the excitatory population activity in the desired range of 3 Hz. However, what might appear to be slightly faster and stronger random fluctuations in the upper three plots of Figure \ref{full_sim_osci}, are in fact regular oscillations across all three depicted variables. While the oscillation amplitudes undergo a rather unpredictable time course, the frequency remains at a constant level of $\simeq 0.5\, Hz$. This feature is also illustrated by the fact that the power spectra depicted in Figure \ref{Power_Spec_without_Analysis} have a peak at the frequency that appeared to be dominant in Figure \ref{full_sim_osci}. The fact that the mean over all power spectra of excitatory thresholds differs from the power spectrum of the mean of these thresholds in its amplitude indicates that not all thresholds oscillate at the same phase. Still though, the overall shape of both spectra are equivalent. 
\begin{figure}
\includegraphics[width=\textwidth]{../../plots/diff_hom/rate_NO_th_compl.png}
%\includegraphics[width=\textwidth]{../../plots/diff_hom/rate_NO_th_closeup.png}
\caption{A/A*: Mean of excitatory (blue) and inhibitory (red) neuronal firing rate. B/B*: NO concentration averaged over readouts at exc. neurons' positions. C/C*: Average exc. firing threshold. Non-diffusive homeostasis was used for $\mathrm{0-750\, s}$ , diffusive h. for $\mathrm{750-1500\, s}$. Target activity was $\mathrm{3\, Hz}$.}
\label{full_sim_osci}
\end{figure}
\begin{figure}
\begin{center}
\includegraphics[width=0.7\textwidth]{../../plots/power_spec/power_spectrum_sim_only.png}
\caption{Blue: Mean of power spectrum over all excitatory thresholds. Red: Power spectrum of mean over all exc. thresholds.}
\label{Power_Spec_without_Analysis}
\end{center}

@@

Describing the neural activity as instantaneous firing rates raises the question how to accurately describe the production of nitric oxide, since the outcome of equations \eqref{Ca_dyn} and \eqref{nNOS_dyn} relies on sudden increases of $Ca^{2+}$ concentration caused by spike events. Naively, one could replace the sum of dirac functions in \eqref{Ca_dyn} by a continuous inflow $Ca^{2+}_{spike} r_i(t)$. If \eqref{nNOS_dyn} was a linear homogeneous differential equation, this approximation would indeed allow for the correct calculation of a linear relation between mean firing rate and NO production. The cubic dependence on $Ca^{2+}$ breaks this simplicity. In order to derive an approximate description, we denote two things: First, the target firing rate of 3 Hz and the corresponding mean interspike interval of 0.33...s is large compared to the decay constant of Calcium, $\tau_{Ca^{2+}} = \mathrm{0.01\,s}$. Consequently, it is very unlikely that one spike event will fall into a region where the Calcium concentration, decaying from the instant jump of the previous spike event, is still significantly larger than zero. As such, one can justify

@@

Returning to the question of how to describe NO synthesis in a simplified form, we note that according to equation \eqref{nNOS_single_sol_approx2}, a single spike causes the release of NO by an amount of $\frac{{Ca^{2+}_{spike}}^3 \tau_{Ca^{2+}}ln(2)}{3}$, which makes the mean rate of NO production over time simply $<nNOS> = <r> \cdot \frac{{Ca^{2+}_{spike}}^3 \tau_{Ca^{2+}}ln(2)}{3}$. However, for two reasons it is not sufficient to simply propose $nNOS_i(t) = r_i(t) \cdot \frac{{Ca^{2+}_{spike}}^3 \tau_{Ca^{2+}}ln(2)}{3}$. First, a single cell fires in the range of $\mathrm{3\,Hz}$, which is - at least on the time scale of the observed oscillations - not enough to assign an "instantaneous" firing rate. Second, the exponential kernel of nNOS leads to a certain amount of delay between the spiking activity and the rate of NO synthesis, see Figure \ref{rate_nNOS_lag}.
\begin{figure}
\begin{center}
\includegraphics[width=0.7\textwidth]{../../plots/diff_hom/rate_nNOS_close.png}
\end{center}
\caption{Average NO synthesis (red) lags behind the mean firing rate of the excitatory population (blue) by $\mathrm{\approx \, 0.1 \,s}$.}
\label{rate_nNOS_lag}
\end{figure}
To deal with this, we introduce an idealized picture of diffusive homeostasis, where diffusion across the tissue is instantaneous. This simplification results in a single level of NO-concentration for all neurons, only being modified over time by the sum of all neurons' NO-syntheses $nNOS_{total}(t) \equiv \sum_{i} nNOS^i(t)$ and the decay term $-\lambda NO$. To account for the spread of NO, $nNOS_{total}$ must be divided by the area $L^2$ of the tissue. Furthermore, we introduce a random fluctuation term $\sigma_{NO} \xi(t)$ that accounts for local, momentary deviations from the mean. The idealized version of \eqref{NO_dyn} then reads
\begin{equation}
\dot{NO}(t)=-\lambda NO + \frac{nNOS_{total}(t)}{L^2} + \sigma_{NO} \xi(t)\;.
\label{NO_dyn_approx}
\end{equation}
The amount of noise in the system is a hard-to-predict quantity since it depends on the spatial structure on the diffusive lattice, giving neurons that are further separated less impact on each other. We tried to estimate the noise level based on the same "instant spread"-simplification made in Equation \eqref{NO_dyn_approx} to describe the overall NO-dynamics. In the final result, this led to an underestimation of the total energy within the power spectrum. However, as long as we assume the noise in the system to be approximately white, noise level only acts as a global scaling factor of the shape of the power spectrum and was thus left open as a model fitting parameter.

From Equation \eqref{NO_dyn_approx} it is clear why we can now describe the \emph{total} influx of NO as the sum of a mean and a random fluctuation: While spiking events on individual cells remain sufficiently separated to justify equation \eqref{nNOS_single_sol_approx2}, the sum of all spike events from the excitatory population results in an effective poisson process with a mean rate of $3 Hz \cdot N_{exc.} = \mathrm{1200\,Hz}$, given the assumption that cells are spiking weakly correlated. Furthermore, we argue that a total rate of $\mathrm{1200\,Hz}$ is sufficiently large to express this rate as a continuous function of time, $r_{total} = N_{exc.}\cdot r_{exc.pop.}(t)$, at least on the time scale of the oscillations to be studied (see  Figure \ref{rate_th_close}). Still though, we need to address the fact that $nNOS_{total}(t)$ lags behind $r_{total}$. We do so by a dynamical instead of a purely functional relationship:
\begin{equation}
\dot{nNOS_{total}}(t) = \frac{r_{exc.pop.}\cdot N_{exc.}\frac{{Ca^{2+}_{spike}}^3 \tau_{Ca^{2+}}ln(2)}{3} - nNOS_{total}}{\tau_{nNOS}}
\label{nNOS_total_dyn}
\end{equation} 
where $\frac{{Ca^{2+}_{spike}}^3 \tau_{Ca^{2+}}ln(2)}{3}$ is the integral from Equation \eqref{nNOS_single_sol_approx1}.

As a next step, we attempted to simplify the dynamics of threshold adaption and its effect on the population rate. Corresponding to the reduction to a single variable $r_{exc.pop.}(t)$ for the mean excitatory activity in equation \eqref{nNOS_total_dyn}, we wanted to find an appropriate description containing only the mean threshold $V_{t,pop.}(t) = <V_t(t)>$. As stated earlier, Figure \ref{rate_th_close} suggests an immediate and approximately linear relation between excitatory population rate and the mean of thresholds within the excitatory population. Figure \ref{thresh_r_linfit} shows a linear fit (being of the form $\alpha x + \beta$) of these two quantities plotted against each other. Apart from rarely appearing high values of $V_{t,pop.}$, we found a good fit to the linear model. We thus expressed the population rate by $r_{exc.pop.}(t) = r_{IP} + \alpha(V_{t,pop.}(t)-V_{t,0})$, where $r_{IP}$ is given in table \ref{Params_IP} as the mean excitatory target firing rate and $V_{t,0} \equiv (r_{IP}-\beta)/\alpha$ is the mean threshold corresponding to the target rate.
\begin{figure}
\includegraphics[width=\textwidth]{../../plots/Spike_Stats/linfit_th_rate.png}
\caption{Linear fit of excitatory population activity and mean excitatory threshold. Slope $\alpha$ of the fit: $\mathrm{-0.789\, Hz/mV}$. Offset $\beta$: $\mathrm{-41.97\, Hz}$. Mean squared error: $\mathrm{R^2 = 0.68}$. Colors encode the time derivative of $<V_t(t)>$, which indicates an ellipsoid like trajectory in phase space, with the firing rate following the threshold with a short delay.}
\label{thresh_r_linfit}
\end{figure}

Combining these results in a set of equations, we find
\begin{align}
\dot{NO}(t) &= -\lambda NO + \frac{nNOS_{total}(t)}{L^2} + \sigma_{NO} \xi(t) \label{2d-sys-homeostasis1}\\
\dot{V_{t,pop.}}(t) &= \frac{NO-NO_0}{NO_0 \tau_{V_t}} \label{2d-sys-homeostasis2} \\
\dot{nNOS_{total}}(t) &= \frac{(r_{IP} + \alpha(V_{t,pop.}(t)-V_{t,0}))\cdot N_{exc.}\frac{{Ca^{2+}_{spike}}^3 \tau_{Ca^{2+}}ln(2)}{3} - nNOS_{total}}{\tau_{nNOS}}\;. \label{2d-sys-homeostasis3}
\end{align}
Through the coordinate transformations
\begin{align}
n &\equiv NO-NO_0 \label{simplif_coord_transf1}\\
\theta &\equiv V_{t,pop.}(t)-V_{t,0} \label{simplif_coord_transf2}\\
nos &\equiv nNOS_{total} - nNOS_{total,0} \label{simplif_coord_transf3}
\end{align}
we can simplify the dynamical system:
\begin{align}
\dot{n} &= -\lambda n + \frac{nos}{L^2} + \sigma_{NO} \xi(t) \label{2d-sys-homeostasis-coord_transf1}\\
\dot{\theta} &= \frac{1}{NO_0 \tau_{V_t}} n \label{2d-sys-homeostasis-coord_transf2} \\
\dot{nos} &= \frac{\alpha \cdot N_{exc} \cdot \gamma \cdot \theta - nos }{\tau_{nNOS}} \label{2d-sys-homeostasis-coord_transf3}
\end{align}
where we have additionally used
\begin{align}
\gamma &\equiv \frac{{Ca^{2+}_{spike}}^3 \tau_{Ca^{2+}}ln(2)}{3} \label{2d-sys-fixed-points1}\\
NO_0 &= \frac{nNOS_{total,0}}{\lambda L^2} \label{2d-sys-fixed-points2}\\
nNOS_{total,0} &= r_{IP}\cdot N_{exc.} \cdot \gamma \; . \label{2d-sys-fixed-points2}
\end{align}
As stated in the beginning, we were interested in finding an analytic expression for the power spectrum of the mean excitatory threshold. Therefore we took the fourier transform of \eqref{2d-sys-homeostasis-coord_transf1} - \eqref{2d-sys-homeostasis-coord_transf3}, yielding
\begin{align}
i\omega f_n &= - \lambda f_n + \frac{f_{nos}}{L^2} + \sigma_{NO} f_{\xi} \label{2d-sys-homeostasis-coord_transf_ft1}\\
i\omega f_\theta &= \frac{f_n}{NO_0\tau_{V_t}} \label{2d-sys-homeostasis-coord_transf_ft2} \\
i\omega f_{nos} &= \frac{\alpha \cdot N_{exc.} \cdot \gamma f_\theta - f_{nos}}{\tau_{nNOS}} \label{2d-sys-homeostasis-coord_transf_ft3}
\end{align}
where $f_{(\cdot)}$ denotes the fourier transform. We solved for $f_\theta$ which immediately gave us the power spectrum $P_\theta (\omega) \equiv |f_\theta (\omega)|^2$ of $\theta$:
\begin{equation}
P_\theta (\omega) = \frac{\sigma_{NO}^2}{\omega^2 \left( \lambda NO_0 \tau_{V_t} + \frac{\alpha N_{exc.} \gamma \tau_{nNOS}}{L^2(\omega^2\tau_{nNOS}^2 + 1)} \right)^2 + \left( \omega^2 NO_0 \tau_{V_t} + \frac{\alpha N_{exc.} \gamma}{L^2 (\omega^2\tau_{nNOS}^2 + 1)}\right)^2} \label{Pow_Spec_Theta_Final}
\end{equation}
Since we assumed the noise term $\xi$ to be approximately white and gaussian noise with unit variance, its power spectrum is $|f_{\xi}|^2 = 1$.

Figure \ref{Pow_Spec_Theta_vs_Analytic} shows the result of \eqref{Pow_Spec_Theta_Final}. 
\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{../../plots/power_spec/power_spectrum_analytic.png}
\end{center}
\caption{Power Spectrum of $\mathrm{\langle V_t \rangle}$ (blue) and analytic predictions based on Equation \eqref{Pow_Spec_Theta_Final}. Red curve was fitted by tweaking only the overall amplitude $\sigma_{NO}$ \eqref{Pow_Spec_Theta_Final} ($\mathrm{\alpha = -0.789 \,Hz/mV}$, as in Figure \ref{thresh_r_linfit}). Green curve is a fit achieved by adjusting $\sigma_{NO}$ and $\alpha$ in \eqref{Pow_Spec_Theta_Final} ($\mathrm{\alpha = -1.114 \,Hz/mV}$).}
\label{Pow_Spec_Theta_vs_Analytic}
\end{figure}
Apparently, the value for $\alpha$ that was determined by means of the linear fit in Figure \ref{thresh_r_linfit} leads to an underestimation of the preferred frequency. Partially, this can be explained by the fact that the linear regression overestimates the importance of a number of high-threshold-outliers which leads to a shallower slope than what would be appropriate for the "working point" of oscillatory behaviour. Another possible explanation is the fact that "instant spread" is only an approximation. Rather, one could picture the nitrous oxide just to fill a fraction of the available space on a short time scale by replacing $L^2$ by a smaller area. This leads to a stronger local increase of concentration, which eventually leads to faster oscillations. Either way, since these parameters always appear as a fraction $\alpha/L^2$ in \eqref{Pow_Spec_Theta_Final}, both explanations have the same effect onto the shape of the power spectrum.

!!! HIER MUSS NOCH EINE ANALYSE HIN, WIE SICH DIE OSCI-AMPLITUDE MIT TAU-TH ÄNDERT !!!

@@

After the elimination of oscillatory activity in the network, we further investigated the features of neural activity, in particular of the excitatory population, since it was exposed to homeostasis.

>>>>

Obviously, the appropriate choice of $NO_0$ is crucial for the goal of achieving and maintaining a certain level of activity. However, one cannot directly set a parameter of the model to the desired population activity, as it was the case for canonical intrinsic homeostasis. Rather, one needs to determine the average concentration \textit{associated} with the desired activity and set it as a target concentration. Though it is possible to derive this relation in an analytic fashion, for practical purposes of the simulation, we let the system run with the previous homeostatic mechanism, still solving equation \eqref{Ca_dyn}-\eqref{NO_dyn} until a steady mean over the concentrations at the neurons' positions was reached. This mean was then set to be the target concentration and we switched to diffusive homeostasis. Table \ref{Params_IP} summarizes the choice of parameters that where introduced in this section.
\begin{table}
\begin{tabular}{|l|l|}
\hline
\textbf{parameter} & \textbf{value} \\
\hline
$\mathrm{r_{IP}}$ & $\mathrm{3\;Hz}$ \\
\hline
$\mathrm{\eta_{IP}}$ & $\mathrm{0.1\;mV}$ \\
\hline
$\mathrm{Ca^{2+}_{spike}}$ & 1 \\ \hline
$\mathrm{\tau_{Ca^{2+}}}$ &  $\mathrm{10\;ms}$ \\
\hline
$\mathrm{\tau_{nNOS}}$ & $\mathrm{100\;ms}$ \\
\hline
$\mathrm{D}$ & default: 12500 $\mathrm{\mu m^2 s^{-1}}$ \\
\hline 
$\mathrm{\lambda}$ & $\mathrm{0.1\;s^{-1}}$ \\
\hline
$\mathrm{\tau_{V_t}}$ & see section \ref{activ_analys} \\
\hline
\end{tabular}
\caption{Parameters of homeostatic intrinsic plasticity.}
\label{Params_IP}
\end{table}

@@

\subsection{Activity Analysis} \label{activ_analys}
As a first attempt, we set the time constant of threshold adaptation in the diffusive homeostasis to $\mathrm{2500\, ms}$, as given by \cite{Sweeney_Paper}. Switching between non-diffusive/diffusive homeostasis after 750 seconds, Figure \ref{full_sim_osci} shows the resulting dynamics of the population activity of the excitatory and inhibitory group, the average $NO$ concentration at the excitatory neurons' positions and the average firing threshold within the exc. group (inh. threshold was fixed, see table \ref{LIF_neuron_params}). Both homeostatic mechanisms managed to keep the excitatory population activity in the desired range of 3 Hz. However, what might appear to be slightly faster and stronger random fluctuations in the upper three plots of Figure \ref{full_sim_osci}, are in fact regular oscillations across all three depicted variables. While the oscillation amplitudes undergo a rather unpredictable time course, the frequency remains at a constant level of $\simeq 0.5\, Hz$. This feature is also illustrated by the fact that the power spectra depicted in Figure \ref{Power_Spec_without_Analysis} have a peak at the frequency that appeared to be dominant in Figure \ref{full_sim_osci}. The fact that the mean over all power spectra of excitatory thresholds differs from the power spectrum of the mean of these thresholds in its amplitude indicates that not all thresholds oscillate at the same phase. Still though, the overall shape of both spectra are equivalent. 
\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{../../plots/diff_hom/rate_NO_th_compl.png}
\end{center}
%\includegraphics[width=\textwidth]{../../plots/diff_hom/rate_NO_th_closeup.png}
\caption{A/A*: Mean of excitatory (blue) and inhibitory (red) neuronal firing rate. B/B*: NO concentration averaged over readouts at exc. neurons' positions. C/C*: Average exc. firing threshold. Non-diffusive homeostasis was used for $\mathrm{0-700\, s}$ , diffusive h. for $\mathrm{700-1500\, s}$. Target activity was $\mathrm{3\, Hz}$.}
\label{full_sim_osci}
\end{figure}
\begin{figure}
\begin{center}
\includegraphics[width=0.7\textwidth]{../../plots/power_spec/power_spectrum_sim_only.png}
\caption{Blue: Mean of power spectrum over all excitatory thresholds. Red: Power spectrum of mean over all exc. thresholds.}
\label{Power_Spec_without_Analysis}
\end{center}

@@

Describing the neural activity as instantaneous firing rates raises the question how to describe the production of nitric oxide, since the outcome of equations \eqref{Ca_dyn} and \eqref{nNOS_dyn} relies on sudden increases of $Ca^{2+}$ concentration caused by spike events. Naively, one could replace the sum of dirac functions in \eqref{Ca_dyn} by a continuous inflow $Ca^{2+}_{spike} r_i(t)$. If \eqref{nNOS_dyn} was a linear homogeneous differential equation, this approximation would indeed allow for the correct calculation of a linear relation between mean firing rate and NO production. The cubic dependence on $Ca^{2+}$ breaks this simplicity. In order to derive an approximate description, we denote two things: First, the target firing rate of 3 Hz and the corresponding mean interspike interval of 0.33...s is large compared to the decay constant of Calcium, $\tau_{Ca^{2+}} = \mathrm{0.01\,s}$. Consequently, it is very unlikely that one spike event will fall into a region where the Calcium concentration, decaying from the instant jump of the previous spike event, is still significantly larger than zero. As such, one can justify

@@

Returning to the question of how to describe NO synthesis in a simplified form, we note that according to equation \eqref{nNOS_single_sol_approx2}, a single spike causes the release of NO by an amount of $\frac{{Ca^{2+}_{spike}}^3 \tau_{Ca^{2+}}ln(2)}{3}$, which makes the mean rate of NO production over time simply $<nNOS> = <r> \cdot \frac{{Ca^{2+}_{spike}}^3 \tau_{Ca^{2+}}ln(2)}{3}$. However, it is not sufficient to simply propose $nNOS_i(t) = r_i(t) \cdot \frac{{Ca^{2+}_{spike}}^3 \tau_{Ca^{2+}}ln(2)}{3}$. A single cell fires in the range of $\mathrm{3\,Hz}$, which is - at least on the time scale of the observed oscillations - not enough to assign an "instantaneous" firing rate.

To deal with this, we introduce an idealized picture of diffusive homeostasis, where diffusion across the tissue is instantaneous. This simplification results in a single level of NO-concentration for all neurons, only being modified over time by the sum of all neurons' NO-syntheses $nNOS_{total}(t) \equiv \sum_{i} nNOS^i(t)$ and the decay term $-\lambda NO$. To account for the spread of NO, $nNOS_{total}$ must be divided by the area $L^2$ of the tissue. Furthermore, we introduce a random fluctuation term $\sigma_{NO} \xi(t)$ that accounts for local, momentary deviations from the mean. The idealized version of \eqref{NO_dyn} then reads
\begin{equation}
\dot{NO}(t)=-\lambda NO + \frac{nNOS_{total}(t)}{L^2} + \sigma_{NO} \xi(t)\;.
\label{NO_dyn_approx}
\end{equation}
The amount of noise in the system is a hard-to-predict quantity since it depends on the spatial structure on the diffusive lattice, giving neurons that are further separated less impact on each other. We tried to estimate the noise level based on the same "instant spread"-simplification made in Equation \eqref{NO_dyn_approx} to describe the overall NO-dynamics. In the final result, this led to an underestimation of the total energy within the power spectrum. However, as long as we assume the noise in the system to be approximately white, noise level only acts as a global scaling factor of the shape of the power spectrum and was thus left open as a model fitting parameter.

From Equation \eqref{NO_dyn_approx} it is clear why we can now describe the \emph{total} influx of NO as the sum of a mean and a random fluctuation: While spiking events on individual cells remain sufficiently separated to justify equation \eqref{nNOS_single_sol_approx2}, the sum of all spike events from the excitatory population results in an effective poisson process with a mean rate of $3 Hz \cdot N_{exc.} = \mathrm{1200\,Hz}$, given the assumption that cells are spiking weakly correlated. Furthermore, we argue that a total rate of $\mathrm{1200\,Hz}$ is sufficiently large to express this rate as a continuous function of time, $r_{total} = N_{exc.}\cdot r_{exc.pop.}(t)$, at least on the time scale of the oscillations to be studied (see  Figure \ref{rate_th_close}). We thus propose
\begin{equation}
nNOS_{total}(t) = r_{exc.pop.}\cdot N_{exc.}\frac{{Ca^{2+}_{spike}}^3 \tau_{Ca^{2+}}ln(2)}{3}
\label{nNOS_total_dyn}
\end{equation} 
where $\frac{{Ca^{2+}_{spike}}^3 \tau_{Ca^{2+}}ln(2)}{3}$ is the integral from Equation \eqref{nNOS_single_sol_approx1}.

As a next step, we wanted to simplify the dynamics of threshold adaption and its effect on the population rate. Corresponding to the reduction to a single variable $r_{exc.pop.}(t)$ for the mean excitatory activity in equation \eqref{nNOS_total_dyn}, we wanted to find an appropriate description containing only the mean threshold $V_{t,pop.}(t) = <V_t(t)>$. As stated earlier, Figure \ref{rate_th_close} suggests an immediate and approximately linear relation between excitatory population rate and the mean of thresholds within the excitatory population. Figure \ref{thresh_r_linfit} shows a linear fit (being of the form $\alpha x + \beta$) of these two quantities plotted against each other. Apart from rarely appearing high values of $V_{t,pop.}$, we found a good fit to the linear model. We thus expressed the population rate by $r_{exc.pop.}(t) = r_{IP} + \alpha(V_{t,pop.}(t)-V_{t,0})$, where $r_{IP}$ is given in table \ref{Params_IP} as the mean excitatory target firing rate and $V_{t,0} \equiv (r_{IP}-\beta)/\alpha$ is the mean threshold corresponding to the target rate.
\begin{figure}
\includegraphics[width=\textwidth]{../../plots/Spike_Stats/linfit_th_rate.png}
\caption{Linear fit of excitatory population activity and mean excitatory threshold. Slope $\alpha$ of the fit: $\mathrm{-0.789\, Hz/mV}$. Offset $\beta$: $\mathrm{-41.97\, Hz}$. Mean squared error: $\mathrm{R^2 = 0.68}$.}
\label{thresh_r_linfit}
\end{figure}

Combining these results in a set of equations, we find
\begin{align}
\dot{NO}(t) &= -\lambda NO + \frac{(r_{IP} + \alpha(V_{t,pop.}(t)-V_{t,0}))\cdot N_{exc.}\frac{{Ca^{2+}_{spike}}^3 \tau_{Ca^{2+}}ln(2)}{3}}{L^2} + \sigma_{NO} \xi(t) \label{2d-sys-homeostasis1}\\
\dot{V_{t,pop.}}(t) &= \frac{NO-NO_0}{NO_0 \tau_{V_t}}\;. \label{2d-sys-homeostasis2}
\end{align}
Through the coordinate transformations
\begin{align}
n &\equiv NO-NO_0 \label{simplif_coord_transf1}\\
\theta &\equiv V_{t,pop.}(t)-V_{t,0} \label{simplif_coord_transf2}\\
\end{align}
we can simplify the dynamical system:
\begin{align}
\dot{n} &= -\lambda n + \frac{\gamma \alpha N_{exc.}}{L^2}\theta + \sigma_{NO} \xi(t) \label{2d-sys-homeostasis-coord_transf1}\\
\dot{\theta} &= \frac{1}{NO_0 \tau_{V_t}} n \label{2d-sys-homeostasis-coord_transf2}
\end{align}
where we have additionally used
\begin{align}
\gamma &\equiv \frac{{Ca^{2+}_{spike}}^3 \tau_{Ca^{2+}}ln(2)}{3} \label{2d-sys-fixed-points1}\\
NO_0 &= \frac{nNOS_{total,0}}{\lambda L^2} \label{2d-sys-fixed-points2}\\
nNOS_{total,0} &= r_{IP}\cdot N_{exc.} \cdot \gamma \; . \label{2d-sys-fixed-points2}
\end{align}
As stated in the beginning, we were interested in finding an analytic expression for the power spectrum of the mean excitatory threshold. Therefore we took the fourier transform of \eqref{2d-sys-homeostasis-coord_transf1} and \eqref{2d-sys-homeostasis-coord_transf2}, yielding
\begin{align}
i\omega f_n &= - \lambda f_n + \frac{\gamma \alpha N_{exc.}}{L^2}f_\theta + \sigma_{NO} f_{\xi} \label{2d-sys-homeostasis-coord_transf_ft1}\\
i\omega f_\theta &= \frac{f_n}{NO_0\tau_{V_t}} \label{2d-sys-homeostasis-coord_transf_ft2} \\
\end{align}
where $f_{(\cdot)}$ denotes the fourier transform. We solved for $f_\theta$ which immediately gave us the power spectrum $P_\theta (\omega) \equiv |f_\theta (\omega)|^2$ of $\theta$:
\begin{equation}
P_\theta (\omega) = \frac{\sigma_{NO}^2}{\omega^2 NO_0^2 \tau_{V_t}^2 \lambda^2 + \left( \omega^2 NO_0 \tau_{V_t} + \frac{\gamma \alpha N_{exc.}}{L^2}\right)^2} \label{Pow_Spec_Theta_Final}
\end{equation}
Since we assumed the noise term $\xi$ to be approximately white and gaussian noise with unit variance, its power spectrum is $|f_{\xi}|^2 = 1$.

Figure \ref{Pow_Spec_Theta_vs_Analytic} shows the result of \eqref{Pow_Spec_Theta_Final}. Choosing noise amplitude to be the only free parameter of the model gave a rather unsatisfying result. Our approximation underestimated the amount of damping within the system (which controls the "peakiness" of the spectrum), as well as predicting a slightly smaller preferred amplitude. While the broadness of the spectrum is mostly controlled by the damping term $-\lambda$, the preferred amplitude is controlled by $\frac{\gamma \alpha N_{exc.}}{L^2}$ and $\frac{1}{NO_0\tau_{V_t}}$. Since the latter term was taken unaltered from the original system (apart from a shift in phase space), we added $-\lambda$ and $\alpha$ as free fitting parameters and managed to get a very good fit to the simulation data, see Figure \ref{Pow_Spec_Theta_vs_Analytic}.   
\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{../../plots/power_spec/power_spectrum_analytic.png}
\end{center}
\caption{Power Spectrum of $\mathrm{\langle V_t \rangle}$ (blue) and analytic predictions based on Equation \eqref{Pow_Spec_Theta_Final}. Red curve was fitted by tweaking only the overall amplitude $\mathrm{\sigma_{NO}}$ in \eqref{Pow_Spec_Theta_Final} ($\mathrm{\alpha = -0.789 \,Hz/mV}$, as in Figure \ref{thresh_r_linfit}). Green curve is a fit achieved by adjusting $\mathrm{\sigma_{NO}}$ and $\mathrm{\lambda}$ in \eqref{Pow_Spec_Theta_Final} ($\mathrm{\lambda = 0.845 \,s^{-1}}$). Black curve was fitted by means of $\mathrm{\sigma_{NO}}$, $\mathrm{\lambda}$ and $\mathrm{\alpha}$ ($\mathrm{\lambda = 0.592 \,s^{-1}}$, $\mathrm{\alpha=-0.989 \,Hz/mV}$).}
\label{Pow_Spec_Theta_vs_Analytic}
\end{figure}
The fact that the value for $\mathrm{\alpha}$ that was determined by means of the linear fit in Figure \ref{thresh_r_linfit} leads to an underestimation of the preferred frequency can be partially explained by the fact that the linear regression overestimates the importance of a number of high-threshold-outliers which leads to a shallower slope than what would be appropriate for the "working point" of oscillatory behaviour. Another possible explanation is the fact that "instant spread" is only an approximation. Rather, one could picture the nitrous oxide just to fill a fraction of the available space on a short time scale by replacing $L^2$ by a smaller area. This leads to a stronger local increase of concentration, which eventually leads to faster oscillations. In either case, since these parameters appear as a fraction $\alpha/L^2$ in \eqref{Pow_Spec_Theta_Final}, both parameters have the same effect onto the shape of the power spectrum.

Regarding the damping of the system, we can interpret   !!! SIMULATION STEHT NOCH AUS!!!

In the context of the presented power spectra, reducing oscillations meant to reduce height and frequency of the peak of the spectrum. We managed to do so by increasing the time constant $\mathrm{\tau_{V_t}}$, as shown in Figure \ref{Peak_Prop_vs_Theta}. Moreover, by choosing $\mathrm{\sigma_{NO}}$, $\mathrm{\lambda}$ and $\mathrm{\alpha}$ in accordance to the fit shown in Figure \ref{Pow_Spec_Theta_vs_Analytic} (black curve), Equation \eqref{Pow_Spec_Theta_Final} provides a good prediction for the position and height of the peak. For proper comparison, we smoothed the curve of the simulation data in plot (A) of Figure \ref{Peak_Prop_vs_Theta} by a window of $\mathrm{[-5,5]}$ data points, since the maximum of the power spectrum of simulation data is likely to overshoot the analytic prediction due to random fluctuations.  
\begin{figure}
\includegraphics[width=\textwidth]{../../plots/diff_hom/osc_ampl_vs_tau_th.png}
\caption{A: Maximum of the spectrum-peak (see Figure \ref{Pow_Spec_Theta_vs_Analytic}). A smoothing window of $\mathrm{[-5,5]}$ data points around the maximum was used to reduce overestimation of the maximum due to random fluctuations. Blue line corresponds to Equation \eqref{Pow_Spec_Theta_Final}, fit values taken from black curve in Figure \ref{Pow_Spec_Theta_vs_Analytic}. B: Position of the peak on the frequency axis. Analytic prediction was acquired according to (A).}
\label{Peak_Prop_vs_Theta}
\end{figure} 
To reduce oscillations as much as possible we eventually settled at a 1000-fold increased time constant, $\mathrm{\tau_{V_t}=2500\,s}$. Despite the slowness of threshold adaption, homeostasis still managed to keep excitatory network activity in the desired range, as shown in Figure \ref{full_sim_osci_slow_tau_th}. Despite a slight regularity of the NO-concentration, no oscillating activity occurs.  
\begin{figure}
\includegraphics[width=\textwidth]{../../plots/diff_hom/rate_NO_th_compl_large_tau.png}
\caption{A: Excitatory (blue) and and inhibitory (red) population rate, being controlled by diffusive homeostasis over the full simulation time. B,C: Mean NO-concentration over excitatory sites and mean excitatory threshold.}
\label{full_sim_osci_slow_tau_th}
\end{figure}
We took this setting as a basis for further analysis of network activity.

@@

After the elimination of oscillatory activity in the network, we further investigated the features of neural activity, in particular of the excitatory population, since it was exposed to homeostasis. This includes statistics of spike timing, as well as of heterogeneous firing rates among the network.
